Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2014-12-17 content/example.tex
add example from Scala library

>>>>

\subsection{An Example from the Scala Collections Library}
The Scala collections library contains the trait \code{TraversableOnce} that is
shared by both collections and iterators over them. The \code{toArray} method of this trait
creates an array and copies the contents of the collection or iterator into it:
%\begin{minted}{scala}
\begin{verbatim}
val result = new Array[B](this.size)
this.copyToArray(result, 0)
\end{verbatim}
%\end{minted}
\otodo{Marianna, can you figure out how to get LaTeX to syntax-highlight this?} 
When \code{this} is an iterator rather than a collection, the call to \code{this.size}
extracts all elements of the iterator to count them. At the call to \code{copyToArray},
the iterator is already empty, so nothing is copied to the newly created array.
One could design an IFDS analysis to detect this kind of bug.

The implementation of \code{TraversableOnce.toArray} is actually correct because the above
code is guarded with a test:
\begin{verbatim}
if(this.isTraversableAgain) ...
\end{verbatim}
When the \code{isTraversableAgain} method returns false, as it does for
an iterator, the \code{toArray} method uses a different
(less efficient) implementation. The bug report would therefore be a false positive. 
The \code{isTraversableAgain} method is easy to analyze: it returns the constant
true in a collection and the constant false in an iterator. However, in order
to eliminate the false positive bug report, an analysis would need to rule out infeasible
paths using correlated calls. Specifically, the following path triggers the bug, but
is infeasible: first, call \code{isTraversableAgain} on a collection, returning
true, then call \code{size} and \code{copyToArray} on an iterator. Our correlated
calls analysis could determine that this path is infeasible because it calls the
collection version of \code{isTraversableAgain} but the iterator versions of
\code{size} and \code{copyToArray}.

\commentout{
    For reference, here is the full code of the method:

def toArray[B >: A : ClassTag]: Array[B] = {
    if (isTraversableAgain) {
        val result = new Array[B](size)
        copyToArray(result, 0)
        result
    }
    else toBuffer.toArray
}

}
